---
title: "Assignment_1"
name: Abhinav Reddy Yelipeddi
date: "2024-02-04"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

dataset source: https://www.kaggle.com/datasets/benroshan/factors-affecting-campus-placement/data

**Importing the dataset**

This R code is using the `readr` package to read a CSV file named "Placement_Data_Full_Class.csv" and it stores the data in a variable called `dataset`.
```{r lib}

library(readr)
dataset<- read_csv("C:/Users/Abhinav Reddy/Desktop/FML/Placement_Data_Full_Class.csv")
```

**Descriptive statistics:**

**Displaying descriptive statistics for quantitative variables**

This R code summarizes and prints descriptive statistics for a specified quantitative variables in a dataset. 

```{r quant}
quantitative_var<-c("sl_no","ssc_p","hsc_p","degree_p","etest_p","mba_p","salary")
quantitative_summary<- summary(dataset[quantitative_var])
print(quantitative_summary)
```
**Displaying descriptive statistics for categorical variables**

This R code shows the frequency distribution of each category within the variables listed in categorical_var.
```{r cate}
categorical_var <- c("gender","ssc_b","hsc_b","degree_t","workex","specialisation")
 categorical_summary<-table(dataset[categorical_var])
print(head( categorical_summary))
```
**Transformation of a variable**

This R code is transforming the salary variable by taking the logarithm and creating a new variable log_transformed_salary.
```{r trans}
dataset$log_transformed_salary <- log(dataset$salary)
head(dataset)
```

This R code is transforming the mba_p variable by taking the square root and creating a new variable sqrt_transformed_mba_p.
```{r trans2}
dataset$sqrt_transformed_mba_p <- sqrt(dataset$mba_p)
head(dataset)
```
**Plotting a quantitative variable**

This R code generates a histogram for the mba_p variable in the dataset.

```{r quan}
quantitative_var1 <- "mba_p"
hist(dataset[[quantitative_var1]], 
     main = "Histogram of mba_p",
     xlab = quantitative_var1,
     col = "lightblue")
```


This R code generates a boxplot for the degree_p variable in the dataset.
```{r boxplot}
boxplot(dataset$degree_p, 
        main = "Box Plot of a degree_p",
        ylab = "Quantitative_Variable",
        col = "lightgreen", border = "black")
```


**Plotting a Scatterplot**

This scatter plot shows the relationship between salary variable and log_transformed_salary variable. 

```{r plot}
plot(x = dataset$'salary', y = dataset$log_transformed_salary, 
     xlab = "salary", ylab = "log_transformed_salary", 
     main = "Scatterplot of Salary vs Transformed Salary")
```




This scatter plot shows the relationship between of ssc_p variable and degree_p variable
```{r plot2}
plot(x = dataset$ssc_p, y = dataset$degree_p,
     xlab = "SSC_Percentage", ylab = "Degree_Percentage",
     main = "Scatter Plot of SSC Percentage vs Degree Percentage")
```

